---
Assignment: "Assignment 10.1"
author: "Zachary DeNoto"
date: August 2 2019
---

Read and clean data
```{r}

library(caret)
library(class)
library(data.table)
library(ggplot2)
library(caTools)


#reads in csv file
cab <- read.csv('cab_ridesv2.csv')

#changes data to data field
cab$Date <- as.Date(cab$Date, format = '%m/%d/%Y')

#looks at dataset structure
str(cab)

#scatterplot to show us distance vs price for all data
ggplot(data = cab, aes(x = cab$distance, y = cab$price)) + geom_point() + 
  ggtitle('Distance vs Price') + labs(x = 'Distance', y = 'Price')
```


Creates and plots smaller sample to better show scatterplot
```{r}
#grabs a random sample to look a less congested scatterplot
cab1 <- data.table(cab)
cab1 <- cab1[sample(.N, 1000)]

#scatterplot of random sample to show us distance vs price better than previous scatterplot
ggplot(data = cab1, aes(x= cab1$distance, y = cab1$price)) + geom_point() + geom_smooth(method = 'lm')  + ggtitle('Distance vs Price for Random Sample of 1000') + labs(x = 'Distance', y = 'Price')

```


Creating subsets of dataset
```{r}
#creates subset for Uber
cab_uber <- cab[which(cab$cab_type =="Uber"),]

#creates subset for Lyft
cab_lyft <- cab[which(cab$cab_type =="Lyft"),]


str(cab_uber)
str(cab_lyft)
  

#grabs a random sample to look a less congested scatterplot
uber1 <- data.table(cab_uber)
uber1 <- uber1[sample(.N, 1000)]

#grabs a random sample to look a less congested scatterplot
lyft1 <- data.table(cab_lyft)
lyft1 <- lyft1[sample(.N, 1000)]
```


Scatterplots of Uber and Lyft datasets
```{r}
#scatterplot of random sample to show us distance vs price better than previous scatterplot
ggplot(data = uber1, aes(x= uber1$distance, y = uber1$price)) + geom_point() + geom_smooth(method = 'lm') + 
  ggtitle('Uber Distance vs Price') + labs(x = 'Distance', y = 'Price')


#scatterplot of random sample to show us distance vs price better than previous scatterplot
ggplot(data = lyft1, aes(x= lyft1$distance, y = lyft1$price)) + geom_point() + geom_smooth(method = 'lm')+ 
  ggtitle('Lyft Distance vs Price') + labs(x = 'Distance', y = 'Price')
```


Correlation tests for Uber and Lyft
```{r}
#testing correlation for price and distance

cor(cab_uber$price, cab_uber$distance, method = 'spearman')

cor(cab_lyft$price, cab_lyft$distance, method = 'spearman')
```


Using linear regression on Uber and Lyft
```{r}
#testing relationship between price and distance
uber_pd <- lm(cab_uber$price ~ cab_uber$distance, data = cab_uber)

lyft_pd <- lm(cab_lyft$price ~ cab_lyft$distance, data = cab_lyft)

summary(uber_pd)
summary(lyft_pd)
```


Making more subsets of Uber and Lyft data
```{r}
#segmenting uber and lyft into 4 categories for weekdays(M-F 7pm), weekends(Friday 8pm -Sunday), days(12am-7pm), and nights(8pm-12am)
weekday_uber <-cab_uber[which(cab_uber$Day == "Monday" |cab_uber$Day == "Tuesday"|cab_uber$Day == "Wednesday" |cab_uber$Day == "Thursday"| (cab_uber$Day == "Friday" & cab_uber$Time <20)),]
weekday_lyft <-cab_lyft[which(cab_lyft$Day == "Monday" |cab_lyft$Day == "Tuesday"|cab_lyft$Day == "Wednesday" |cab_lyft$Day == "Thursday"| (cab_lyft$Day == "Friday" & cab_lyft$Time <20)),]
  
weekend_uber <-cab_uber[which(cab_uber$Day == "Saturday" |cab_uber$Day == "Sunday"|(cab_uber$Day == "Friday" & cab_uber$Time > 19)),]
weekend_lyft <-cab_lyft[which(cab_lyft$Day == "Saturday" |cab_lyft$Day == "Sunday"|(cab_lyft$Day == "Friday" & cab_lyft$Time > 19)),]
  
  
day_uber <- cab_uber[which(cab_uber$Time < 20),]
day_lyft <- cab_lyft[which(cab_lyft$Time < 20),]
  
night_uber <- cab_uber[which(cab_uber$Time > 19),]
night_lyft <- cab_lyft[which(cab_lyft$Time > 19),]
```


Using linear and multiple linear regression to test out differences
```{r}
#testing relationship between price and distance for weekdays
uber_wkday <- lm(weekday_uber$price ~ weekday_uber$distance, data = weekday_uber)
lyft_wkday <- lm(weekday_lyft$price ~ weekday_lyft$distance, data = weekday_lyft)

summary(uber_wkday)
summary(lyft_wkday)

#testing relationship between price and distance for weekends
uber_wkend <- lm(weekend_uber$price ~ weekend_uber$distance, data = weekend_uber)
lyft_wkend <- lm(weekend_lyft$price ~ weekend_lyft$distance, data = weekend_lyft)

summary(uber_wkend)
summary(lyft_wkend)

#testing relationship between price and distance for day time
uber_daytime <- lm(day_uber$price ~ day_uber$distance, data = day_uber)
lyft_daytime <- lm(day_lyft$price ~ day_lyft$distance, data = day_lyft)

summary(uber_daytime)
summary(lyft_daytime)

#testing relationship between price and distance for night time
uber_nightime <- lm(night_uber$price ~ night_uber$distance, data = night_uber)
lyft_nightime <- lm(night_lyft$price ~ night_lyft$distance,   data = night_lyft)

summary(uber_nightime)
summary(lyft_nightime)
```


Using multiple linear regression to test out differences
```{r}
#testing relationship between price, distance and type of ride for weekdays
uber_wkday <- lm(weekday_uber$price ~ weekday_uber$distance + factor(weekday_uber$name), data = weekday_uber)
lyft_wkday <- lm(weekday_lyft$price ~ weekday_lyft$distance + factor(weekday_lyft$name), data = weekday_lyft)

summary(uber_wkday)
summary(lyft_wkday)

#testing relationship between price, distance and type of ride  for weekends
uber_wkend <- lm(weekend_uber$price ~ weekend_uber$distance + factor(weekend_uber$name) , data = weekend_uber)
lyft_wkend <- lm(weekend_lyft$price ~ weekend_lyft$distance + factor(weekend_lyft$name) , data = weekend_lyft)

summary(uber_wkend)
summary(lyft_wkend)

#testing relationship between price, distance and type of ride  for day time
uber_daytime <- lm(day_uber$price ~ day_uber$distance + factor(day_uber$name) , data = day_uber)
lyft_daytime <- lm(day_lyft$price ~ day_lyft$distance + factor(day_lyft$name)  ,data = day_lyft)

summary(uber_daytime)
summary(lyft_daytime)

#testing relationship between price, distance and type of ride for night time
uber_nightime <- lm(night_uber$price ~ night_uber$distance + factor(night_uber$name), data = night_uber)
lyft_nightime <- lm(night_lyft$price ~ night_lyft$distance + factor(night_lyft$name),   data = night_lyft)

summary(uber_nightime)
summary(lyft_nightime)




#testing relationship between price, distance and type of ride for weekdays

summary(lm(weekday_uber$price ~ weekday_uber$distance + factor(weekday_uber$name)+ weekday_uber$surge_multiplier, data = weekday_uber))
summary(lm(weekday_lyft$price ~ weekday_lyft$distance + factor(weekday_lyft$name)+ weekday_lyft$surge_multiplier, data = weekday_lyft))

#testing relationship between price, distance and type of ride  for weekends

summary(lm(weekend_uber$price ~ weekend_uber$distance + factor(weekend_uber$name)+ weekend_uber$surge_multiplier , data = weekend_uber))
summary(lm(weekend_lyft$price ~ weekend_lyft$distance + factor(weekend_lyft$name)+ weekend_lyft$surge_multiplier , data = weekend_lyft))

#testing relationship between price, distance and type of ride  for day time

summary(lm(day_uber$price ~ day_uber$distance + factor(day_uber$name)+ day_uber$surge_multiplier , data = day_uber))
summary(lm(day_lyft$price ~ day_lyft$distance + factor(day_lyft$name)+ day_lyft$surge_multiplier  ,data = day_lyft))

#testing relationship between price, distance and type of ride for night time

summary(lm(night_uber$price ~ night_uber$distance + factor(night_uber$name)+ night_uber$surge_multiplier, data = night_uber))
summary(lm(night_lyft$price ~ night_lyft$distance+ factor(night_lyft$name)+ night_lyft$surge_multiplier,   data = night_lyft))
```


Plots surges for Lyft 
```{r}
#shows when surges occur for lyft
ggplot(data = cab_lyft, aes(x = cab_lyft$Time, y = cab_lyft$surge_multiplier)) + geom_point() + 
  ggtitle('Lyft Time vs Surge') + labs(x = 'Time', y = 'Surge Multiplier')


#shows when surges occur on weekends for lyft
ggplot(data = weekend_lyft, aes(x = weekend_lyft$Time, y = weekend_lyft$surge_multiplier)) + geom_point()+ 
  ggtitle('Lyft Time vs Surge for Weekends') + labs(x = 'Time', y = 'Surge Multiplier')

#shows when surges occur on weekdays for lyft
ggplot(data = weekday_lyft, aes(x = weekday_lyft$Time, y = weekday_lyft$surge_multiplier)) + geom_point()+ 
  ggtitle('Lyft Time vs Surge for Weekdays') + labs(x = 'Time', y = 'Surge Multiplier')
```


Creating subsets of Uber and Lyft by day of the week
```{r}
#creates a Monday subset for lyft
mon_lyft <- cab_lyft[which(cab_lyft$Day =="Monday"),]
#creates a Tuesday subset for lyft
tues_lyft <- cab_lyft[which(cab_lyft$Day =="Tuesday"),]
#creates a Wednesday subset for lyft
wed_lyft <- cab_lyft[which(cab_lyft$Day =="Wednesday"),]
#creates a Thursday subset for lyft
thur_lyft <- cab_lyft[which(cab_lyft$Day =="Thursday"),]
#creates a Friday subset for lyft
fri_lyft <- cab_lyft[which(cab_lyft$Day =="Friday"),]
#creates a Saturday subset for lyft
sat_lyft <- cab_lyft[which(cab_lyft$Day =="Saturday"),]
#creates a Sunday subset for lyft
sun_lyft <- cab_lyft[which(cab_lyft$Day =="Sunday"),]
```


Plotting surges for Lyft by day of the week
```{r}
#shows when surges occur on Monday for lyft
ggplot(data = mon_lyft, aes(x = mon_lyft$Time, y = mon_lyft$surge_multiplier)) + geom_point() + 
  ggtitle('Lyft Time vs Surge for Monday') + labs(x = 'Time', y = 'Surge Multiplier')

#shows when surges occur on Tuesday for lyft
ggplot(data = tues_lyft, aes(x = tues_lyft$Time, y = tues_lyft$surge_multiplier)) + geom_point()+ 
  ggtitle('Lyft Time vs Surge for Tuesday') + labs(x = 'Time', y = 'Surge Multiplier')

#shows when surges occur on Wednesday for lyft
ggplot(data = wed_lyft, aes(x = wed_lyft$Time, y = wed_lyft$surge_multiplier)) + geom_point()+ 
  ggtitle('Lyft Time vs Surge for Wednesday') + labs(x = 'Time', y = 'Surge Multiplier')

#shows when surges occur on Thursday for lyft
ggplot(data = thur_lyft, aes(x = thur_lyft$Time, y = thur_lyft$surge_multiplier)) + geom_point()+ 
  ggtitle('Lyft Time vs Surge for Thursday') + labs(x = 'Time', y = 'Surge Multiplier')

#shows when surges occur on Friday for lyft
ggplot(data = fri_lyft, aes(x = fri_lyft$Time, y = fri_lyft$surge_multiplier)) + geom_point()+ 
  ggtitle('Lyft Time vs Surge for Friday') + labs(x = 'Time', y = 'Surge Multiplier')

#shows when surges occur on Saturday for lyft
ggplot(data = sat_lyft, aes(x = sat_lyft$Time, y = sat_lyft$surge_multiplier)) + geom_point()+ 
  ggtitle('Lyft Time vs Surge for Saturday') + labs(x = 'Time', y = 'Surge Multiplier')

#shows when surges occur on Sunday for lyft
ggplot(data = sun_lyft, aes(x = sun_lyft$Time, y = sun_lyft$surge_multiplier)) + geom_point()+ 
  ggtitle('Lyft Time vs Surge for Sunday') + labs(x = 'Time', y = 'Surge Multiplier')

```


Plotting Price and Date for Uber and Lyft
```{r}

#plots price and date for uber
ggplot(cab_uber, aes(x = cab_uber$Date, y = cab_uber$price)) +
  geom_point() + ggtitle('Date vs Price for Uber') + labs(x = 'Date', y = 'Price')


#plots price and date for lyft
ggplot(cab_lyft, aes(x = cab_lyft$Date, y = cab_lyft$price)) +
  geom_point() + ggtitle('Date vs Price for Lyft') + labs(x = 'Date', y = 'Price')


```


Predictions
```{r}
split <- sample.split(cab, SplitRatio = 0.8)
split

train <- subset(cab, split == "Uber")
test <- subset(cab, split == "Lyft")


u_or_l <- glm(cab$cab_type ~ cab$price + cab$name + cab$Time + cab$distance, data = cab, family = 'binomial')

summary(u_or_l)


anova(u_or_l,test="Chisq")

res <- predict(u_or_l, test, type = 'response')
res


res <- predict(u_or_l, train, type = 'response')
res

confmatrix <- table(Actual_value = cab$cab_type, Predicted_value = res > 0.5)
confmatrix

confmatrix[[1,1]] /sum(confmatrix)


```


Do not run this...it will crash the program. It was an attempt to use k-nearest neighbor on dataset
```{r}
#ccab <- read.csv('cab_ridesv3.csv')

# 
# dat.d <- sample(1:nrow(ccab),size=nrow(ccab)*0.7,replace = FALSE) #random selection of 70% data.
# 
# train2 <- ccab[dat.d,] # 70% training data
# test2 <- ccab[-dat.d,] # remaining 30% test data
# 
# bin_train_labels <- ccab[dat.d,1]
# bin_test_labels <-ccab[-dat.d,1]
# 
# 
# bin_test <- knn(train = train2, test = test2,cl = bin_train_labels, k=100)
# 
# CrossTable(x = bin_test_labels, y = bin_test, prop.chisq = FALSE)


```






